---
title: "Homework 6"
author: "Yuki Joyama"
date: "2023-11-19"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read packages
```{r message=FALSE}
library(tidyverse)
```

# Problem 1
```{r message=FALSE}
# read csv file
df_homicide = read_csv("./data/homicide-data.csv")

# modify the dataframe
df_homicide = df_homicide |> 
  mutate(
    city_state = paste(city, state, sep = ", "), # create a city_state variable
    victim_age = as.numeric(victim_age), # change victim_age from char to numeric
    homicide = if_else(
      disposition == "Closed without arrest" | disposition == "Open/No arrest", "unsolved","resolved"
    ),
    homicide = as.factor(homicide)
  ) |>
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"))) |> # omit some cities due to the lack of data or data entry mistake
  filter(victim_race %in% c("White", "Black")) |> # limiting data to white or black
  filter(victim_sex != "Unknown") |> # exclude unknown sex
  drop_na(victim_age) # exclude unknown age
```

Logistic regression for crime outcome and victim age, sex and race in Baltimore, MD.

```{r}
# run logistic regression for Baltimore, MD
log_baltimore = df_homicide |> 
  filter(city_state == "Baltimore, MD") |> 
  glm(homicide ~ victim_age + victim_sex + victim_race, family = binomial, data = _) |> 
  broom::tidy()

log_baltimore

# check the level of the outcome variable
levels(pull(df_homicide, homicide))

# odds ratio for coef and 95% CI 
log_baltimore |> 
  filter(term == "victim_sexMale") |> 
  mutate(
    or = exp(estimate), # calculate odds ratio of the coefficients
    lb.or = exp(estimate - qnorm(.975) * std.error), # obtain the lower level of 95% CI
    ub.or = exp(estimate + qnorm(.975) * std.error) # obtain the upper level of 95% CI
  ) |> 
  select(or, lb.or, ub.or) |> 
  knitr::kable(digits = 2)
```

In Baltimore, MD, male victims have approximately 2.35 times the odds (95% CI: 1.79 - 3.08) of resolved cases than the unresolved cases compared to female victims when all other variables fixed.

Now run the logistic regression for each of the cities.
```{r}
# set up a function to get odds ratio for coef and 95% CI from logistic regression output
or_ci_glm = function(df){
  df |> 
    filter(term == "victim_sexMale") |> 
    mutate(
      or = exp(estimate),
      lb.or = exp(estimate - qnorm(.975) * std.error),
      ub.or = exp(estimate + qnorm(.975) * std.error)
    ) |> 
    select(or, lb.or, ub.or) 
}


case_log = df_homicide |>
  nest(df = -city_state) |> 
  mutate(
    log_model = purrr::map(df, \(df) glm(homicide ~ victim_age + victim_sex + victim_race, family = binomial, data = df)), # save logistic models for each city as list
    results = map(log_model, broom::tidy) # save logistic model results in tibble as list
  ) |>
  mutate(
    ors = map(results, or_ci_glm) # save odds ratios and 95%CI as list
  ) |> 
  unnest(ors) |> 
  select(city_state, or, lb.or, ub.or) 
  
# visualize the results
case_log |> 
  mutate(city_state = fct_reorder(city_state, or)) |> 
  ggplot(aes(x = city_state, y = or)) +
  geom_point() +
  geom_errorbar(aes(x = city_state, ymin = lb.or, ymax = ub.or)) +
  labs(
    title = "The odds ratios of resolved cases among male compared to female",
    x = "City",
    y = "Odds ratio"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) 
```

From the plot, we can see that in Albuquerque, NM, the odds of cases to be solved is the lowest in male compared to female holding other variables (OR: `r case_log |> filter(city_state == "Albuquerque, NM") |> select(or) |> round(2)`, 95% CI: `r case_log |> filter(city_state == "Albuquerque, NM") |> select(lb.or) |> round(2)` - `r case_log |> filter(city_state == "Albuquerque, NM") |> select(ub.or) |> round(2)`). In New York, NY, the odds of resolved cases in male compared to female was the highest keeping all other variables fixed (OR: `r case_log |> filter(city_state == "New York, NY") |> select(or) |> round(2)`, 95% CI: `r case_log |> filter(city_state == "New York, NY") |> select(lb.or) |> round(2)` - `r case_log |> filter(city_state == "New York, NY") |> select(ub.or) |> round(2)`).

# Problem 2

```{r}
# download dataset for this question
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```








